### 2020-02-04 presentations/jsps_2020-03-01/index.Rmd
Add Rmd to generate slides for JSPS conference talk

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---
class: inverse, middle, title-slide, big-margin
<!-- background-position: 50% 50%
background-image: url(images/title_slide_background.svg)
background-size: contain -->

# 日本のシダ植物多様性の多次元解析
<br>
## 新田ジョエル^1^、海老原淳^2^

^1^スミソニアン国立自然史博物館植物研究部<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

^2^国立科学博物館植物研究部

日本植物分類学会第19回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg")
```

```{r load-packages, cache = FALSE}
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("code/functions.R"))

# Load targets
loadd(list = c(
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "ja_bioregions_ferns",
  "biodiv_results_pteridos",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

```

---
class: big-margin

## Biodiversity is in danger from climate change

---
class: big-margin

## We must understand the distribution of biodiversity to save it

- Most analyses of BD only include richness
- But BD is more complicated: evo history also important

---
class: big-margin

## Japan is an ideal site to study fern BD

- Extremely well sampled

---
class: big-margin

## Goals

- Map biodiversity (species richness, PD, endemism)
- Identify biogeographical realms
- Understand drivers of biogeo structure

---
class: big-margin

## Data

- Occurrence data of Japanese fern specimens
- Phylogenetic tree (*rbcL*)
- Morphological traits
- Climate data

---

## Species richness peaks on Yakushima Isl.

```{r richness-plot, out.width = "120%", fig.asp = 0.7}
alpha_div_ferns <-
  alpha_div_ferns %>%
  filter(latitude > 21)

alpha_div_ferns_binned <-
  alpha_div_ferns %>%
  mutate(
    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
      str_split(",") %>%
      map_chr(1) %>%
      parse_number) %>%
  group_by(latitude) %>%
  summarize_at(
    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
  )

# Fern richness
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_scico(palette = "bamako", na.value="grey") +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

b <-
  ggplot(alpha_div_ferns_binned, aes(y = latitude, x = richness)) +
  geom_path() +
  labs(
    x = "No. species"
  ) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))

```

---

## Phylogenetic diversity is high in N and S

```{r mpd-plot, out.width = "120%", fig.asp = 0.7}

# pterido_phy_mpd
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "phy_mpd_obs_z", 
  metric_title = "MPD",
  label = "Mean Phylogenetic Distance") +
  scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

b <- ggplot(
  alpha_div_ferns_binned, 
  aes(y = latitude, x = phy_mpd_obs_z)) +
  geom_path() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    position = "right") +
  labs(
    subtitle = "",
    x = "MPD"
  ) + 
  annotate(
    "segment", 
    x = 0, xend = 0, 
    y = -Inf, yend = Inf, 
    linetype = 2) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))

```

---

## Highest bioregional diversity in S islands

```{r mapequation-prep, include = FALSE}
# Convert all_cells dataframe to sf object
all_cells_sf <-
  purrr::map2(all_cells$longitude, all_cells$latitude, ~st_point(c(.x, .y))) %>%
  st_sfc(crs = 4326) %>%
  st_sf(all_cells, .)

# Join secondary grid cells with bioregions
all_cells_regions <-sf::st_join(
  all_cells_sf,
  ja_bioregions_ferns) %>% 
  select(secondary_grid_code, longitude, latitude, bioregion) %>% 
  as_tibble %>%
  mutate(bioregion = as.factor(bioregion)) %>%
  filter(secondary_grid_code %in% occ_data_ferns$secondary_grid_code)
```

```{r mapequation-plot, out.width = "120%", fig.asp = 0.7}
ggplot(all_cells_regions, aes(x = longitude, y = latitude, fill = bioregion)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Bioregion") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )

```

---

## Honshu has significant ecological structure

```{r ecostructure}
# For easy manipulation of plot size
tweak = 0.5

# Output pie plot without basemap
ecos_plot_pie2(
  omega =  species_motifs_ferns_6L$omega,
  image_width = 6000*tweak,
  image_height = 4800*tweak,
  radius = 0.05,
  no_map = TRUE,
  blank_bg = "white",
  long_lim = c(122, 146),
  lat_lim = c(24, 46),
  path = here::here("presentations/jsps_2020-03-01/images/species_motifs_ferns_6L.png")
)

# crop image, save as species_motifs_ferns_6L_cropped.png
```

.center[<img src="images/species_motifs_ferns_6L_cropped.png" alt="drawing" height="600">]

---

## S islands have high rates of endemism

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
ggplot(biodiv_results_pteridos, aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1", na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

---

## Neoendemic areas have more sexual diploids

```{r canape-sex, out.width = "120%", fig.asp = 0.7}

percent_sex_dip_ferns_round <-
  percent_sex_dip_ferns %>% 
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3))

canape_sex_mode <-
  biodiv_results_pteridos %>%
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3)) %>%
  inner_join(
    percent_sex_dip_ferns_round,
    by = c("latitude", "longitude")
  ) %>%
  assert(is_uniq, secondary_grid_code) %>%
  filter(!is.na(canape))

# Run analysis of variance (AOV) on
# latitudinal breadth by reproductive mode.
canape_by_repro_model = aov(
  percent_sex_dip ~ canape,
  data = canape_sex_mode)

canape_by_repro_model_summary = tidy(canape_by_repro_model)

# AOV of latidudinal breadth by reproductive mode
# showed a significant difference, so
# run Tukey HSD test on results.
canape_by_repro_tukey = TukeyHSD(canape_by_repro_model) %>% tidy

ggplot(canape_sex_mode, aes(x = canape, y = percent_sex_dip)) +
  geom_jitter() +
  geom_boxplot() +
  coord_flip()

```