### 2020-06-24 reports/figures/figures.Rmd
Add working report to generate figures

---
title: "figures"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)

# Load packages and functions
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
library(patchwork)
library(rcartocolor)

# Load objects from cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

loadd(list = c(
  "ses_div_ferns",
  "shape_ferns",
  "comm_scaled_coverage_0.2",
  "ses_div_ferns_endemic",
  "k_taxonomy",
  "k_phylogeny",
  "regions_taxonomy",
  "regions_phylogeny",
  "species_motifs_ferns"
),
cache = ja_fern_cache)
```

```{r plot-prep}
tax_regions_data <- regions_taxonomy %>%
  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

phy_regions_data <- regions_phylogeny %>%
  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

ses_div_ferns_spatial <-
  shape_ferns %>%
  left_join(categorize_endemism(ses_div_ferns), by = c(grids = "site")) %>%
  left_join(comm_scaled_coverage_0.2, by = c(grids = "site")) %>%
  mutate(redundancy = 1 - (richness/abundance)) %>%
  left_join(tax_regions_data, by = "grids") %>%
  left_join(phy_regions_data, by = "grids")

ses_div_ferns_endemic_spatial <-
shape_ferns %>%
  left_join(categorize_endemism(ses_div_ferns_endemic), by = c(grids = "site"))

map_theme <- theme(
    # panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "white", size = 0.2),
    axis.ticks = element_blank()
  )

```

```{r k-taxonomy-plot, include = FALSE}
k_taxonomy$df %>%
  as_tibble() %>%
  mutate(optimal = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE)) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal)) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    y = "Explained variances",
    x = "Number of clusters") +
  annotate("text", x = Inf, y = Inf, label = glue::glue("K = {k_taxonomy$optimal$k}"), hjust = 1, vjust = 1) +
  theme(legend.position = "none")
```

```{r k-phylogeny-plot, include = FALSE}
k_phylogeny$df %>%
  as_tibble() %>%
  mutate(optimal = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE)) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal)) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    y = "Explained variances",
    x = "Number of clusters") +
  annotate("text", x = Inf, y = Inf, label = glue::glue("K = {k_phylogeny$optimal$k}"), hjust = 1, vjust = 1) +
  theme(legend.position = "none")
```

Redundancy (number of specimens per species) is generally high throughout the country, but lower in Hokkaido and some coastal areas and small islands. Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas. Raw MPD is uniform throughout, but SES of MPD shows clustering in mountainous areas of Honshu and overdispersion in souther Kyushu and islands further south.

```{r diversity-plots, fig.width = 8, fig.height = 8}
# Richness
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() +
  labs(fill = "Richness \n (n spp.)") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# Sampling rendundancy
b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Redundancy") +
  map_theme +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

# MPD
c <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs)) +
  geom_sf(color = "transparent") +
  labs(fill = "MPD") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  scale_fill_viridis_c() +
  map_theme

# SES MPD
d <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES MPD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

a + b + c + d  +
  plot_annotation(tag_levels = 'A')

```

**Figure 1**. Taxonomic and phylogenetic diversity of the ferns of Japan. A, richness. B, sampling redundancy. C, mean phylogenetic distance (MPD). D, Standard effect size of MPD.

```{r endem-plots, fig.width = 8, fig.height = 10}
# PE
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  map_theme

# SES PE
b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pe_obs_z, "abs")
    )) +
  map_theme

# PE (endemic)
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  map_theme

# SES PE (endemic)
d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
    )) +
  map_theme

# CANAPE
e <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  map_theme

# CANAPE (endemic)
f <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  map_theme

a + b + c + d + e + f +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')
```

**Figure 2**. Phylogenetic endemism of the ferns of Japan. A, phylogenetic endemism (PE). B, Standard effect size (SES) of PE. C, PE of endemic taxa only. D, SES of PE of endemic taxa only. E, Categorical analysis of neo- and paleo- endemism (CANAPE) results. F, CANAPE results for endemic taxa only.

```{r bioregion-plots, fig.width = 8}
# Taxonomic bioregions
a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
  map_theme

# Phylogenetic bioregions
b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
  geom_sf(color = "transparent") +
  labs(fill = "Cluster") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
  map_theme

a + b +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')
```

**Figure 3**. Regionalization of the ferns of Japan. A, taxonomic regionalization. B, phylogenetic regionalization.

```{r ecostructure}
tweak <- 0.75

sites <- ses_div_ferns_spatial %>%
  pull(grids)

centroids <-
ses_div_ferns_spatial %>%
  select(geometry) %>%
  mutate(centroid = sf::st_centroid(geometry)) %>%
  sf::st_coordinates(x = .$centroid) %>%
  as_tibble() %>%
  mutate(site = sites)

omega <-
species_motifs_ferns$omega %>%
  as.data.frame %>%
  rownames_to_column("site") %>%
  left_join(centroids, by = "site") %>%
  column_to_rownames("site")

omega_centroids <- select(omega, X, Y) %>% # longitude, latitude
  as.matrix()

omega <- select(omega, -X, -Y)

ecos_plot_pie2(
    omega = omega,
    image_width = 6000*tweak,
    image_height = 4800*tweak,
    radius = 0.08,
    no_map = TRUE,
    coords = omega_centroids,
    long_lim = c(121, 148),
    lat_lim = c(23, 47),
    color = carto_pal(8, "Safe"),
    blank_bg = "grey10",
    path = here::here(
      "reports/figures/species_motifs_ferns.png"
    )
)

```

![](species_motifs_ferns.png)

**Figure 3**. Ecological structure of the ferns of Japan. K = 8.