### 2020-07-10 presentations/bsa_2020-07-27/index.Rmd
Add slides for BSA talk

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg")
```

```{r load-packages, cache = FALSE}
library(magick)
# library(gt)
library(ggtext)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(rcartocolor)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("R/functions.R"))

# Load targets
loadd(list = c(
  "ses_div_ferns_spatial",
  "occ_point_data_ferns"),
  cache = ja_fern_cache)

```

```{r title-slide-image, eval = FALSE}
# Code to make title slide image, don't need to run when rendering slides
title_theme <- theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none",
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

# Change cluster to factor
ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  title_theme

b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPD") +
  scale_fill_scico(
    palette = "roma",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  title_theme

c <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
  geom_sf(color = "transparent") +
  labs(fill = "Cluster") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  title_theme

a + b + c + plot_layout(ncol = 1)

# Write out here, then crop manually
ggsave(
  plot = title_image, 
  filename = here::here("presentations/bsa_2020-07-27/images/title.png"),
  dpi = 300)

```

class: inverse, middle, title-slide, big-margin
background-color: #000000
background-position: 50% 50%
background-image: url(images/title_background.png)
background-size: contain

.pull-left-wide[
# Exploring dimensions of biodiversity in Japanese ferns
<br>
## Joel H. Nitta<sup>1</sup>, Atsushi Ebihara<sup>2</sup>

<sup>1</sup>University of Tokyo<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

<sup>2</sup>National Museum of Nature and Science, Japan

Botanical Society of America<br><span style = 'font-size: 80%;'>2020.07.27</span>
]

.pull-right-narrow[
![](images/title_cropped.png)
]

---
class: big-margin

## Biodiversity is a two-sided coin

Characterizing the **distribution of biodiversity** is needed to:

- understand the **processes** generating it

- understand how to **conserve** it

---
class: big-margin

## Most studies of biodiversity count species

However, taxonomic richness is only one dimension of biodiversity!

We should also consider

- Phylogenetic diversity

- Functional diversity

- How these are related to each other

This requires comprehensive sampling of taxa, DNA, and traits

---
class: big-margin

## Ferns of Japan

.pull-left[

- One of the best-collected fern floras in the world

- Extensive study by both scientists and amateur societies

]

.pull-right[
```{r gbif-colls-per-country}
# Read in pteridophyte GBIF collections per country data from
# Iwasaki lab presentation
collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
  mutate(country = factor(country)) %>%
  mutate(country = fct_reorder(country, n))

# Plot showing top 10
ggplot(collections_by_country, aes(x = country, y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "GBIF pteridophyte \ncollections (top 10)",
    y = "No. collections") +
  scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(colour="black"),
      axis.text.y = ggplot2::element_text(colour="black")
  )
```
]

---
class: big-margin

## Ferns of Japan

.pull-left[
- One of the best-collected fern floras in the world

- Extensive study by both scientists and amateur societies
]

.pull-right[
```{r ja-pterido-coll-curve}
# Format fern abundances (no hybrids) as numeric vector
fern_abun <-
  occ_point_data_ferns %>%
  group_by(taxon) %>%
  count(sort = TRUE) %>%
  pull(n)

# Run iNEXT
inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")

# create subset of observed points to add to plot as dots
observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]

# check estimated completeness
est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")

# Make iNEXT plot
inext_out$iNextEst %>%
  filter(method != "observed") %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=3) +
  scale_y_continuous("Richness") +
  scale_x_continuous("No. specimens") +
  labs(
    title = "Species collection curve",
    subtitle = "Japanese ferns (no hybrids)"
  ) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(colour="black"),
    axis.text.y = ggplot2::element_text(colour="black"),
    legend.position="bottom", 
    legend.title=element_blank()
    )

```
]

---
class: big-margin

## Goals of this study

Using the **ferns of Japan** as a study system, seek to understand:

1. How is biodiversity distributed?

2. How is biodiversity structured?

3. How well is biodiversity protected?