### 2021-06-03 ms/data_exploration.Rmd
Add Rmd with data exploration for modeling

---
title: "Data exploration"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  knit_type: manual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we perform initial data exploration to determine the appropriate models to fit to the data.

```{r load-packages, include = params$knit_type == "manual"}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)

# Load custom functions
source(here::here("R/functions.R"))
```

There are two data sets:
* `biodiv_ferns_cent_env` with biodiversity metrics measured on the full set of species (species missing functional data were excluded when measuring functional diversity metrics).
* `biodiv_ferns_cent_repro` with biodiveristy metrics related to genetic diversity measured only on species with reproductive mode data available.

All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`).

The `cent` part of the name indicates that they include geographic locations as centroids of each grid cell.

Below, "Full dataset" refers to `biodiv_ferns_cent_env` and "Repro dataset" refers to `biodiv_ferns_cent_repro`.

```{r load-data}
# Load the data from the targets workflow
tar_load(biodiv_ferns_cent_env)
tar_load(biodiv_ferns_cent_repro)
```

### Structure of the data

```{r struc-full-data}
biodiv_ferns_cent_env
```

```{r struc-repro-data}
biodiv_ferns_cent_repro
```

The columns are as follows:

* `grids`: ID code for each grid cell
* `lat`: latitude of grid cell centroid
* `long`: longitude of grid cell centroid
* `richness`: number of taxa
* `pd_obs_z`: standard effect size (SES) of phylogenetic diversity (pd)
* `fd_obs_z`: SES of functional diversity (fd)
* `rpd_obs_z`: SES of relative pd
* `rfd_obs_z`: SES of relative fd
* `temp`: mean annual temperature (10 x °C)
* `precip`: total annual precipitation (mm)
* `precip_season`: seasonality in precipitation
* `percent_apo`: percent of apogamous taxa in that grid cell (repro data only)

### Check for missing data

```{r check-missing}
# These will return TRUE if all data are non-missing, 
# or an error if any data are missing (NA)
biodiv_ferns_cent_env %>%
  assert(not_na, everything(), success_fun = success_logical)

biodiv_ferns_cent_repro %>%
  assert(not_na, everything(), success_fun = success_logical)
```

No missing data.

### Check for outliers

We will first check for outliers in the data by visualizing box plots. Dots at the extremes may be outliers.

```{r box-plots-full}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  labs(title = "Full dataset")
```

```{r box-plots-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  labs(title = "Repro dataset")
```

There are a few points beyond the whiskers that may be outliers.

Next, check for outliers using Grubb's test. Significant p-values indicate that the most extreme value is an outlier.

```{r grubbs-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

```{r grubbs-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

So the one percent apomictic > 60% appears to be an outlier. Perhaps this has a very small number of species, most of which happen to be apogamous?

```{r check-apo-outlier}
biodiv_ferns_cent_repro %>%
  slice_max(1, order_by = percent_apo) %>%
  left_join(select(biodiv_ferns_cent_env, grids, richness)) %>%
  select(grids:long, percent_apo, richness)
```

Looks like that is the case---just three species in this grid cell! This data point should be excluded. Let's do that and check for outliers again.

```{r remove-apo-outlier-check-grubbs}
biodiv_ferns_cent_repro %>% 
  filter(percent_apo != max(percent_apo)) %>%
  pull(percent_apo) %>%
  grubbs_tidy()
```

OK, no more outliers in `percent_apo`.

### Normality

Check for normality with histograms and QQ-plots.

Inspect histograms.

```{r hist-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Full dataset")
```

```{r hist-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Repro dataset")
```

Inspect qq-plots.

```{r qqplot-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Full dataset")
```

The response variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, `rpd_obs_z` all look decent. `richness` is skewed, but it is count data and we aren't using a Gaussian model for that anyways (poisson or negative binomial), so that's OK.

```{r qqplot-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Repro dataset")
```

Again, the response variables `pd_obs_z`, and `rpd_obs_z` look OK.

### Check for zeros in count data

Zero-inflated data can be a problem for linear models.

```{r check-for-zeros}
biodiv_ferns_cent_env %>%
  mutate(rich_zero = richness == 0) %>%
  arrange(richness) %>%
  summarize(
    total_zero = sum(rich_zero),
    perc_zero = sum(rich_zero) / nrow(.))
```

No zeros in our count data (richness).

### Plot raw data

Here we plot response variables (row labels) vs. indep variables (column labels) for each dataset, and add a trend line.

```{r plot-raw-data-env, fig.height = 8}
ggplot(biodiv_ferns_cent_env, aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(temp, precip, precip_season),
    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z)) +
  labs(title = "Full dataset")
```

`richness` vs. `temp` definitely shows a hump-shaped trend line, and the other response variables vs. `temp` also show one if not quite as strongly. The other relationships appear to be linear or uncorrelated.

```{r plot-raw-data-repro}
# Remove the outlier first
biodiv_ferns_cent_repro %>% 
  filter(percent_apo != max(percent_apo)) %>%
ggplot(aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(percent_apo, temp, precip, precip_season),
    rows = vars(pd_obs_z, rpd_obs_z)) +
  labs(title = "Repro dataset")
```

As one would expect from the last plot, `temp` shows a (negative) hump shape. However, there aren't many points on the right.. `percent_apo` seems more linear, especially for `rpd_obs_z`, but it's subtle. Anyways, we will probably keep the quadratic term for temperature but use a linear model for `percent_apo`.

### Check for overdispersion in richness

Since richness is count data (only takes positive values), we will use either a poisson or negative binomial to model its distribution. The poisson distribution requires that the ratio of residual variance to degrees of freedom be approximately 1; significantly different ratios are "overdispersed" or "underdispersed". Check for this with a dispersion test.

```{r check-overdispersion}
# Create general linear model with poisson distribution
richness_poisson_mod <- glm(
  richness ~ temp + I(temp^2) + precip + precip_season, 
  data = biodiv_ferns_cent_env, family = poisson)

# Check ratio of residual variance to df (should be near 1 to meet poisson assumptions)
richness_poisson_mod$deviance / richness_poisson_mod$df.residual

# Run statistical test
AER::dispersiontest(richness_poisson_mod, trafo=1)
```

That's very overdispersed. So we won't use poisson, but negative binomial instead.