### 2023-04-17 ms/shida-no-kai-figs.Rmd
Add Rmd to produce images with Japanese

---
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions:
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output:
  bookdown::word_document2:
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
    toc: no
    number_sections: no
    reference_docx: american-journal-of-botany.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
params:
  doc_type: doc
editor_options:
  chunk_output_type: console
---

---
bibliography:
  - `r here::here("ms/references.yaml")`
  - `r here::here("ms/references_other.bib")`
---

```{r manual, include = FALSE, evaluate = FALSE}
# Run these lines if rendering outside of targets plan
library(targets)
library(tarchetypes)
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
```

```{r setup, include = FALSE, cache = FALSE}
# Load libraries used for rendering Rmds
library(flextable)
library(kableExtra)
library(ftExtra)
library(scales)
library(patchwork)
library(RColorBrewer)
library(ggspatial) # for scalebars on maps

# Set knitr options
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "hide", cache = FALSE)

# Load functions and variables only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)
```

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  aic_env_repro,
  binary_sig_results,
  biodiv_ferns_cent,
  biodiv_ferns_spatial,
  broad_alignment_list,
  comm_ferns,
  comm_ferns_full,
  csl_file,
  deer_range,
  fern_traits,
  green_list,
  japan_fern_tree,
  japan_map_points,
  japan_pterido_tree,
  japan_rbcL,
  japan_shp,
  k_phylogeny,
  k_taxonomy,
  lat_span_summary,
  lrt_results,
  model_fits,
  model_params,
  ms_functions,
  occ_point_data_summary,
  phy_sig_results,
  plastome_calibration_dates,
  plastome_tree,
  ppgi,
  protected_areas,
  refs_other_yaml,
  refs_yaml,
  refs_other_bib,
  repro_data,
  signif_cells_protected_area,
  signif_cells_deer_danger,
  template_file,
  traits_summary,
  biodiv_ferns_spatial_lumped,
  bioregion_cutoff
))
```

```{r plot-prep}
# (color palettes are set in ms_functions.R)

# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *p*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

indep_var_lookup <- tribble(
  ~term, ~term_pretty,
  "temp", "Temperature",
  "precip", "Precipitation",
  "precip_season", "Precipitation seasonality",
  "I(temp^2)", "Temperature^2^",
  "percent_apo", "% apomictic taxa"
)

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial)) %>%
  st_transform(4612)

# Format biodiv_ferns_spatial data for plotting
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(
      endem_type,
      levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all
    # variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x,
      "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  ) %>%
  st_transform(4612)

fig_dpi <- 600

# For powerpoint presentations, increase base font size
theme_set(theme_grey(base_size = 20))
```

```{r raw-diversity-make}
trend_line_theme <- function () {
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )
}

# Convert diversity metrics dataframe to centroids for plotting trends
ses_div_centroids <-
  biodiv_ferns_spatial %>%
  mutate(
    longitude = sf::st_centroid(geom) %>% st_coordinates %>% magrittr::extract(,1),
    latitude = sf::st_centroid(geom) %>% st_coordinates %>% magrittr::extract(,2)) %>%
  as_tibble %>%
  dplyr::select(-geom)
# Get min and max latitude for setting common limits between
# maps and trendlines
ja_lat_min <- ses_div_centroids %>% pull(latitude) %>% min()
ja_lat_max <- ses_div_centroids %>% pull(latitude) %>% max()

# Fern richness
# - map
a <-
  ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  geom_sf() +
  scale_fill_scico(palette = "lajolla") +
  labs(fill = "種数") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(
    legend.title = element_text(size = 30/.pt),
    legend.text = element_text(size = 24/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )
# - latitudinal trends
b <- ggplot(ses_div_centroids, aes(x = latitude, y = richness)) +
  geom_point(color = "grey50", shape = 1 ) +
  geom_smooth(se = FALSE) +
  xlim(ja_lat_min, ja_lat_max) +
  coord_flip() +
  labs(y = "種数") +
  trend_line_theme()
a + b + plot_layout(widths = c(3,1), heights = c(1,1))

ggsave("working/shida-no-kai/richness.png")
```

```{r rand-diversity-make}
# PD
pd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") + #nolint
  # label
  labs(title = "PD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/pd_rand.png")

# FD
fd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") + #nolint
  # label
  labs(title = "FD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/fd_rand.png")

# RPD
rpd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") + #nolint
  labs(title = "RPD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/rpd_rand.png")
```

```{r bioregions-make}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(title = "分類学的距離") +
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  map_theme()

# Phylogenetic bioregions
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = phylo_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(title = "系統的距離") +
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(axis.text.y = element_blank())

bioregions_plot <- a + b +
  plot_layout(ncol = 2)

ggsave("working/shida-no-kai/bioregions.png")

# Taxonomic, plain version
ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  map_theme()

ggsave("working/shida-no-kai/bioregions_taxonomic.png")
```

```{r find-top-species}
# Extract clusters
clusters <-
  biodiv_ferns_spatial_lumped %>%
  as_tibble() %>%
  select(site = "grids", cluster = taxonomic_cluster, -geom)
# Convert community df to long format
comm_ferns_long <-
  comm_ferns %>%
  rownames_to_column("site") %>%
  pivot_longer(names_to = "taxon", values_to = "abun", -site)
# Find top-scoring species (with high occurrence inside one cluster, and low elsewhere)
top_species <-
  comm_ferns_long %>%
  left_join(clusters, by = "site") %>%
  filter(cluster %in% 1:4) %>%
  group_by(taxon, cluster) %>%
  summarize(
    abun = sum(abun),
    total = n()
  ) %>%
  mutate(
    # percent abundance within a bioregion
    perc_abun = abun / total
  ) %>%
  arrange(taxon, desc(perc_abun)) %>%
  select(taxon, cluster, perc_abun) %>%
  group_by(taxon) %>%
  # take top two highest percent abundances per species, calculate their difference
  slice_max(n = 2, order_by = perc_abun, with_ties = FALSE) %>%
  mutate(
    top_cluster = cluster[perc_abun == max(perc_abun)]
  ) %>%
  summarize(
    max_dif = max(perc_abun) - min(perc_abun),
    cluster = unique(top_cluster)
  ) %>%
  # keep only those with greater than 65% difference between top 1st and 2nd 
  # most abundant bioregions
  filter(max_dif > 0.5) %>%
  group_by(cluster) %>%
  # Choose top three per cluster for plotting
  slice_max(n = 3, order_by = max_dif, with_ties = FALSE)
```

```{r plot-top-species-prep}
# make helper function to plot top-scoring species
plot_top_taxon <- function (taxon_select, ses_div_ferns_spatial, comm_ferns_long) {
ses_div_ferns_spatial %>%
  left_join(
    filter(comm_ferns_long, taxon == taxon_select),
    by = c(grids = "site")
  ) %>%
  mutate(abun = factor(abun)) %>%
  ggplot(aes(fill = abun)) +
  scale_fill_manual(values = c("#404040", "#d01c8b")) +
  geom_sf(color = "transparent") +
  labs(
    subtitle = str_replace_all(taxon_select, "_", " ")
  ) +
  map_theme() +
  theme(
    plot.subtitle = element_text(face = "italic"),
    legend.position = "none"
  )
}
```

```{r plot-top-species-cluster-1, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 1, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 1") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[1]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[2]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[3]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_1.png")
```

```{r plot-top-species-cluster-2, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 2, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 2") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[4]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[5]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[6]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_2.png")
```

```{r plot-top-species-cluster-3, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 3, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 3") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[7]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[8]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[9]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_3.png")
```

```{r plot-top-species-cluster-4, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 4, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 4") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[10]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[11]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[12]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_4.png")
```

```{r rpd-vs-apo}
# Extract dataset used for model, including significance columns
biodiv_ferns_used_for_model <- biodiv_ferns_spatial %>%
  # drop geometry (only need centroids for modeling)
  st_set_geometry(NULL) %>%
  as_tibble() %>%
  # filter to only sites used in model
  inner_join(
    biodiv_ferns_cent %>% as_tibble() %>% select(grids),
    by = "grids"
  )

rpd_vs_apo_data <-
  model_fits %>%
  mutate(
    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
    resp_var_print = case_when(
      resp_var == "richness" ~ "Richness (no. spp.)",
      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
    ),
    model_data = list(biodiv_ferns_used_for_model),
    signif_cols = list(signif_cols)) %>%
  filter(resp_var == "rpd_obs_z", indep_var == "percent_apo")

ggplot(rpd_vs_apo_data$model_data[[1]], aes(x = percent_apo, y = rpd_obs_z)) +
  labs(
    y = "RPDの効果量",
    x = "無配生殖種の割合") +
  geom_point(aes(color = rpd_signif), shape = 16, size = 3) +
  scale_color_manual(values = signif_cols) +
  scale_x_continuous(labels = function(x) paste0(round(x * 100), "%")) +
  geom_line(data = rpd_vs_apo_data$fit[[1]], alpha = 0.35, color = "blue") +
  geom_ribbon(
    data = rpd_vs_apo_data$fit[[1]],
    aes(ymin = low, ymax = up), 
    alpha = 0.15
  ) +
  theme(legend.position = "none")

ggsave("working/shida-no-kai/rpd_vs_apo.png")
```

```{r protected-areas, out.width = "120%", fig.asp = 0.65, dev="png", dpi=200}
japan_box <- sf::st_bbox(biodiv_ferns_spatial)

protection_cols_ja_map <- protection_cols
names(protection_cols_ja_map) <- str_to_lower(names(protection_cols_ja_map))
protection_cols_ja_map <- protection_cols_ja_map[c("medium", "high")]

status_labels <-
protected_areas %>%
  # area is in sq m, so convert to sq kilom first
  mutate(area = st_area(.) / 1e6) %>%
  mutate(
    area = as.numeric(area),
    area_chr = area %>% round %>% scales::number(big.mark = ","),
    area_percent = area %>% magrittr::divide_by(377900) %>% scales::percent(),
    label = glue::glue("{str_to_sentence(status)} ({area_percent})") %>% set_names(status)
) %>%
  pull(label) %>%
  str_replace_all("High", "高") %>%
  str_replace_all("Medium", "中") %>%
  set_names(c("high", "medium"))

ggplot() +
  geom_sf(data = japan) +
  geom_sf(data = protected_areas, aes(fill = status), color = "transparent") +
  coord_sf(xlim = c(japan_box$xmin, japan_box$xmax), ylim = c(japan_box$ymin, japan_box$ymax)) +
  labs(
    fill = "保全レベル"
  ) +
  scale_fill_manual(values = protection_cols_ja_map, labels = status_labels[names(protection_cols_ja_map)]) +
  theme(
    axis.ticks = element_blank()
  ) + 
  labs(
    caption = "環境省\nhttps://www.biodic.go.jp/biodiversity/activity/policy/map/map17/"
  )

ggsave("working/shida-no-kai/cons_areas.png")
```

```{r conserv-stats}
# Convert signif_cells_protected_area to wide format
signif_cells_protected_area_wide <- #nolint
  signif_cells_protected_area %>%
  select(status, metric, percent_protected) %>%
  pivot_wider(names_from = status, values_from = percent_protected) %>%
  rowwise %>%
  mutate(total = sum(medium, high)) %>%
  ungroup

# Calculate min and max percent area protected overall
# - define helper functions
calc_min_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    min() %>%
    scales::percent(accuracy = 0.1)
}
calc_max_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    max() %>%
    scales::percent(accuracy = 0.1)
}

min_protected_med <- calc_min_pa(signif_cells_protected_area_wide, medium)
max_protected_med <- calc_max_pa(signif_cells_protected_area_wide, medium)
min_protected_high <- calc_min_pa(signif_cells_protected_area_wide, high)
max_protected_high <- calc_max_pa(signif_cells_protected_area_wide, high)
min_protected_total <- calc_min_pa(signif_cells_protected_area_wide, total)
max_protected_total <- calc_max_pa(signif_cells_protected_area_wide, total)

# Calculate min and max protected area for the best protected category
signif_cells_protected_area_best <- # nolint
  signif_cells_protected_area_wide %>%
  slice_max(total)

# Verify that the best-protected biodiv type is PE
assertthat::assert_that(signif_cells_protected_area_best$metric == "pe")

best_protected_med <- signif_cells_protected_area_best %>%
  pull(medium) %>%
  scales::percent(accuracy = 0.1)
best_protected_high <- signif_cells_protected_area_best %>%
  pull(high) %>%
  scales::percent(accuracy = 0.1)
best_protected_total <- signif_cells_protected_area_best %>%
  pull(total) %>%
  scales::percent(accuracy = 0.1)

# Calculate the total area of Japan
area_of_japan <-
  japan_shp %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  pull(area)

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_protected <-
  protected_areas %>%
  st_make_valid %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  filter(status %in% c("medium", "high")) %>%
  group_by(status) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_protected = total_area / area_of_japan) %>%
  mutate(percent_protected = as.numeric(percent_protected)) %>%
  # Add sum for high + medium
  bind_rows(
    summarize(
      .,
      percent_protected = sum(percent_protected),
      total_area = sum(total_area)
    ) %>% mutate(status = "total")
  )

# Extract values for medium and high protection level
total_percent_protected_medium <- total_percent_protected %>%
  filter(status == "medium") %>%
  pull(percent_protected)
total_percent_protected_high <- total_percent_protected %>%
  filter(status == "high") %>%
  pull(percent_protected)
total_percent_protected_all <- sum(
  total_percent_protected_medium, total_percent_protected_high)

### Deer ###

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_deer_danger <-
  deer_range %>%
  mutate(
    area = st_area(.) %>% units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  group_by(range) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_danger = total_area / area_of_japan) %>%
  mutate(percent_danger = as.numeric(percent_danger))

# Convert to a list for easy indexing
deer_overall <- total_percent_deer_danger %>%
  mutate(
    # Format numbers for nice printing
    area = total_area %>%
      as.numeric %>%
      number(accuracy = 1) %>%
      paste("km^2^"),
    percent = scales::percent(percent_danger, accuracy = 0.1)
  ) %>%
  split(.$range)

deer_in_high_biodiv <-
signif_cells_deer_danger %>%
  group_by(range) %>%
  summarize(percent_danger = mean(percent_danger), .groups = "drop") %>%
  mutate(
    # Format numbers for nice printing
    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
  split(.$range)

deer_highest_danger <- signif_cells_deer_danger %>%
  slice_max(percent_danger) %>%
  # Make sure biodiv type with greatest danger is richness in estimated dataset
  verify(metric == "richness") %>%
  verify(range == "estimated") %>%
  mutate(
    # Format numbers for nice printing
    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
  split(.$range)
```

```{r conserv-status-make}
# Add total protected area (medium + high)
signif_cells_protected_area_total <- #nolint
  signif_cells_protected_area %>%
  group_by(metric) %>%
  summarize(percent_protected = sum(percent_protected), .groups = "drop") %>%
  mutate(status = "total")

total_percent_protected_for_plot <- #nolint
  total_percent_protected %>%
  select(status, percent_protected) %>%
  mutate(status = str_to_sentence(status))

# Plot protected areas
signif_cells_protected_area %>%
  select(metric, status, percent_protected) %>%
  bind_rows(signif_cells_protected_area_total) %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "種数",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "種数"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(
    data = total_percent_protected_for_plot,
    aes(xintercept = percent_protected, linetype = status),
    color = "grey30"
  ) +
  scale_linetype_manual(
    values = protection_lines,
    labels = c("中", "高", "合計"),
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_fill_manual(
    values = protection_cols,
    labels = c("中", "高", "合計"),
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_x_continuous(
    labels = function(x) scales::percent(x, accuracy = 1),
    expand = expansion(mult = c(0, .05))) +
  labs(
    x = "保全されている割合",
    y = "生物多様性の指標",
    fill = "保全レベル",
    linetype = "保全レベル"
  ) +
  theme_bw(base_size = 24) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "right"
  )

ggsave("working/shida-no-kai/protection_levels.png")

names(deer_lines) <- str_replace_all(
  names(deer_lines),
  "Estimated",
  "予想"
)

names(deer_cols) <- str_replace_all(
  names(deer_cols),
  "Estimated",
  "予想"
)

# Plot areas in danger from herbivory by deer
signif_cells_deer_danger %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "種数",
    ),
    range = str_to_sentence(range) %>%
      str_replace_all("Estimated", "予想")
  ) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "種数"))) %>%
  ggplot(aes(x = percent_danger, y = metric, fill = range)) +
  geom_col(position = position_dodge()) +
  geom_vline(
    data = total_percent_deer_danger %>% mutate(
      range = str_to_sentence(range) %>%
      str_replace_all("Estimated", "予想")),
    aes(linetype = range, xintercept = percent_danger),
    color = "grey30"
  ) +
  scale_linetype_manual(
    values = deer_lines,
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_fill_manual(
    values = deer_cols,
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_x_continuous(
    labels = function(x) scales::percent(x, accuracy = 1),
    expand = expansion(mult = c(0, .05))) +
  labs(
    x = "シカの面積",
    y = "生物多様性の指標",
    fill = "時代",
    linetype = "時代"
  ) +
  theme_bw(base_size = 20) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right"
  )

ggsave("working/shida-no-kai/deer_levels.png")
```

```{r conserv-status-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("conserv-status"), "png"),
  conserv_status_plot, width = 7.25, height = 4, units = "in", dpi = fig_dpi)
```

```{r conserv-status-print, fig.width = 7.25, fig.height = 4, eval = params$doc_type == "pdf"}
# Print plot if rendering to PDF
conserv_status_plot
```

**`r figure("conserv-status") %>% str_to_upper()`**. Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan. (A) Percentage of land area with medium or high protected status for grid cells with significantly high biodiversity by protection status. (B) Percentage of land area occupied by deer for grid cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data). Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods). For (A), vertical lines indicate total percentage area protected across Japan (baseline protection rate). For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).